Smart Drill-Down

### 2015-02-17 TableSummarization/Paper/algorithms.tex
Changes to paper

Note that the reader may be wondering why we did not simply count the score of each rule using a variant of the a-priori algorithm in one pass, and then pick the set of rules that maximizes score subsequently. This is because doing so will lead to a sub-optimal set of rules: by not accounting for the rules that have already been selected, we will not be able to ascertain the marginal benefit of adding an additional rule correctly.

>>>>

The reader may be wondering why we did not simply count the score of each rule using a variant of the a-priori algorithm in one pass, and then pick the set of rules that maximizes score subsequently. This is because doing so will lead to a sub-optimal set of rules: by not accounting for the rules that have already been selected, we will not be able to ascertain the marginal benefit of adding an additional rule correctly.

### 2015-02-17 TableSummarization/Paper/exp.tex
Changes to paper

\begin{figure}
\vspace{-5pt}
\centering
\includegraphics[width=90mm]{graphs/screenshots6col/empty_expansion.jpg}
\caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
\end{figure}

@@

\subsection{Qualitative Study}
We first perform a qualitative study of smart drill down. We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset), and then try out different weight functions to study their effects.
\subsubsection{User interface testing}
We now present the rule-based summaries displayed as a result of a few different user actions.

@@

% TODO: Add the experiment where you compare to vanilla a priori.
\vspace{-3pt}
\subsection{Quantitative Study}\label{sec:quantitative_study}
The performance of our algorithm depends on various parameters, such as $m_w$ (the max weight) and $minSS$ (minimum required sample size). In this section, we study the effects of these parameters on the computation time and accuracy of our algorithm. We use both the Marketing and Census datasets. The Marketing dataset is relatively small with around $9000$ tuples, whereas the Census dataset is quite large, with $2.5$ million tuples. We observe that the accuracy and running time of our algorithm depends largely on $m_w$ and $minSS$, rather than the underlying database size.

>>>>

\begin{figure}
\vspace{-5pt}
\centering
\includegraphics[width=90mm]{graphs/screenshots6col/empty_expansion.jpg}
\caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
\vspace{-15pt}
\end{figure}

@@

\subsection{Qualitative Study}\label{sec:exp_qualitative}
We first perform a qualitative study of smart drill down. We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset), and then try out different weight functions to study their effects.
\subsubsection{User interface testing}
We now present the rule-based summaries displayed as a result of a few different user actions.

@@

% TODO: Add the experiment where you compare to vanilla a priori.
\vspace{-3pt}
\subsection{Quantitative Study}\label{sec:quantitative_study}
The performance of our algorithm depends on various parameters, such as $m_w$ (the max weight) and $minSS$ (minimum required sample size). In this section, we study the effects of these parameters on the computation time and accuracy of our algorithm. We use both the Marketing and Census datasets. The Marketing dataset is relatively small with around $9000$ tuples, whereas the Census dataset is quite large, with $2.5$ million tuples. The accuracy of our algorithm depends on $m_w$ and $minSS$, rather than the underlying database size. The worst case running time for large datasets is close to the time taken for making one pass on the dataset. When we expand a rule using an existing sample in memory, the running time is small and only depends on $minSS$ rather than on the dataset size.

### 2015-02-17 TableSummarization/Paper/formal.tex
Changes to paper

\stitle{Size Weighting Function:} 
\techreporttext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
Here we set weight equal to the number of non-starred values in the rule $r$; we define this quantity to be the {\em size} of the rule. Consider the examples in Table \ref{table:sizescoringexample}. The weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$. Thus, the total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$. If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then the $MCount$ of rule ($a$, $\star$) would reduce to $600$, since $300$ of its tuples would be covered by the previous rule $(a, b_2)$. Thus, our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$. If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$. Thus, when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).}

\papertext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
Here we set weight equal to the number of non-starred values in the rule $r$ i.e. the {\em size} of the rule. For example, in Table~\ref{table:introexample3}, the rule (Target, bicycles, $\star$) has weight $2$. To get the idea behind using this weighting function, imagine we are trying to reconstruct the table from the rules. We start with a blank value in every column of every table tuple. Then for each displayed rule $r$, we fill each of its values in the corresponding column of $MCount(r)$ tuples. We do this for each rule, filling a total of $\sum_{r} MCount(r)W(r)$ values. Thus maximising score leads to choosing the set of rules that will pre-fill as much of the table as possible.}

>>>>

\stitle{Size Weighting Function:} 
\techreporttext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
Here we set weight equal to the number of non-starred values in the rule $r$; we define this quantity to be the {\em size} of the rule. Consider the examples in Table \ref{table:sizescoringexample}. The weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$. Thus, the total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$. If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then the $MCount$ of rule ($a$, $\star$) would reduce to $600$, since $300$ of its tuples would be covered by the previous rule $(a, b_2)$. Thus, our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$. If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$. Thus, when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).} \papertext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
Here we set weight equal to the number of non-starred values in the rule $r$ i.e. the {\em size} of the rule. For example, in Table~\ref{table:introexample3}, the rule (Target, bicycles, $\star$) has weight $2$. To get the idea behind using this weighting function, imagine we are trying to reconstruct the table from the rules. We start with a blank value in every column of every table tuple. Then for each displayed rule $r$, we fill each of its values in the corresponding column of $MCount(r)$ tuples. We do this for each rule, filling a total of $\sum_{r} MCount(r)W(r)$ values. Thus maximising score leads to choosing the set of rules that will pre-fill as much of the table as possible.}

### 2015-02-17 TableSummarization/Paper/intro.tex
Changes to paper

To reiterate, our chief contribution in this paper is the {\em smart drill down} interaction operator,
an extension of traditional drill down, aimed at allowing analysts to zoom into the
more ``interesting'' parts of a dataset. In addition to this operator,
we develop techniques to support this operator on increasingly larger datasets:
\squishlist
\item {\em Basic Interaction:} We demonstrate that finding the optimal list of rules is {\sc NP-Hard},
and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation. 
\item {\em Dynamic Sample Maintenance:} To improve response time on large tables, we formalize the problem of dynamically
maintaining samples in memory to support smart drill down. 
We show that optimal identification of samples is once again {\sc NP-Hard}, and 
we develop an approximate scheme for dynamically maintaining and using multiple samples of the table in memory.
\squishend

>>>>

To reiterate, our chief contribution in this paper is the {\em smart drill down} interaction operator,
an extension of traditional drill down, aimed at allowing analysts to zoom into the
more ``interesting'' parts of a dataset. In addition to this operator,
we develop techniques to support this operator on increasingly larger datasets:
\squishlist
\item {\em Basic Interaction:} We demonstrate that finding the optimal list of rules is {\sc NP-Hard},
and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation. 
\item {\em Dynamic Sample Maintenance:} To improve response time on large tables, we formalize the problem of dynamically
maintaining samples in memory to support smart drill down. 
We show that optimal identification of samples is once again {\sc NP-Hard}, and 
we develop an approximate scheme for dynamically maintaining and using multiple samples of the table in memory.
\squishend
We have a prototype of the smart drill down operator that is completely implemented and usable. Starting from this point on, whenever we provide result snippets, these will be screenshots from our prototype. In Section~\ref{sec:exp_qualitative}, we will compare smart drill-down with traditional drill-down and show that smart drill-down returns considerably better results; as a matter of fact, smart drill-down can be viewed as a generalization of traditional drill-down (when the weighting function is set to equal $1$ if the drill-down column has a non-$\star$ value and $0$ otherwise), so our implemented prototype also supports traditional drill-down.

### 2015-02-17 TableSummarization/Paper/related.tex
Changes to paper

There is work on constructing `explanation tables', which are sets of rules that co-occur with a given binary attribute of the table~\cite{DBLP:journals/pvldb/GebalyAGKS14}. This work again focuses on displaying rules that will cause the resulting max entropy distribution to best approximate the actual distribution of values.

>>>>

There is work on constructing `explanation tables', which are sets of rules that co-occur with a given binary attribute of the table~\cite{DBLP:journals/pvldb/GebalyAGKS14}. This work again focuses on displaying rules that will cause the resulting max entropy distribution to best approximate the actual distribution of values. A few vision papers~\cite{data-deluge, cetintemel2013query} suggest frameworks for building interactive data exploration systems. Some of these ideas, like maintaining user profiles, could be integrated into a smart drill down system. Reference~\cite{DBLP:conf/edbt/CandanCQS09} proposes an extension to OLAP drill-down that takes visualization real estate into account, by clustering attribute values. But it focuses on expanding a single column at a time, and relies on a given value hierarchy for clustering.

### 2015-02-17 TableSummarization/Paper/sampling.tex
Changes to paper

In Section~\ref{sec:sampling_algorithms}, we describe the problem of optimally allocating memory to different samples. We show that the problem is NP-Hard, and mention an approximate solution using an additional assumption. In Section~\ref{sec:sample-handler}, we describe a component of the system called the {\em SampleHandler}, which is responsible for creating and maintaining multiple samples of different parts of the table in memory, and creating temporary samples for BRS to process. Detailed descriptions of the sampling problem and the SampleHandler are in the technical report. The technical report also contains some additional optimizations, and ways to set the minimum sample size.

@@

When a user attempts to drill down on $r$, the SampleHandler takes all $n_r$ tuples from $s_r$, and also tuples covered by $r$ from samples $s_{r'}$ for all $r' \in U$ that are sub-rules of $r$. If the total number of such tuples is $\geq minSS$, then we run BRS on that set of tuples to perform the drill down. If not, then we need to access the hard disk to obtain more tuples covered by $r$.

@@

\papertext{We present the idea behind the approximate solution below. A more detailed description can be found in the technical report. The solution consists of two main steps:
\squishlist
\item \textbf{Locally Optimal Solutions :} For each rule $r_0 \in U \setminus L$, let $M_{r_0}$ be the set consisting of $r_0$ and its leaf children. We consider the number of tuples assigned to members of $M_{r_0}$. Each such assignment of numbers has a `cost' equal to total number of tuples assigned, and `value' equal to sum of probabilities of $M_{r_0}$ members with $ess > minSS$. We consider all `locally optimal' assignments, where probability value cannot be increased without increasing the cost.
\item \textbf{Combining Solutions :} Then, we use dynamic programming to pick one optimal solution per $r_0$ so as to get maximum probability value with the given cost constraint. We discretize costs, say to have granularity $100$. Let $M_{0}, M_{1}, ... M_{D}$ be the sets for which we found locally optimal solutions. We make $D$ iterations in all, where in the $i^{th}$ iteration, we try different locally optimal solutions for $M_i$ and use it to update the maximum attainable value for each cost. 
\squishend
}

>>>>

In Section~\ref{sec:sampling_algorithms}, we describe the problem of optimally allocating memory to different samples. We show that the problem is NP-Hard, and describe an approximate solution. In Section~\ref{sec:sample-handler}, we describe a component of the system called the {\em SampleHandler}, which is responsible for creating and maintaining multiple samples of different parts of the table in memory, and creating temporary samples for BRS to process. Detailed descriptions of the sampling problem and the SampleHandler are in the technical report. The technical report also contains some additional optimizations, and ways to set the minimum sample size.

@@

When a user attempts to drill down on $r$, the SampleHandler takes all $n_r$ tuples from $s_r$, and also tuples covered by $r$ from samples $s_{r'}$ for all $r' \in U$ that are sub-rules of $r$. If the total number of such tuples is $\geq minSS$, then we run BRS on that set of tuples to perform the drill down. Note that this set of tuples forms a uniformly random sample of tuples covered by $r$. If not, then we need to access the hard disk to obtain more tuples covered by $r$.

@@

\papertext{We present the idea behind the approximate solution below. A more detailed description can be found in the technical report. The solution consists of two main steps:
\squishlist
\item \textbf{Locally Optimal Solutions :} For each rule $r_0 \in U \setminus L$, let $M_{r_0}$ be the set consisting of $r_0$ and its leaf children. We consider the number of tuples assigned to members of $M_{r_0}$. Each such assignment of numbers has a `cost' equal to total number of tuples assigned, and `value' equal to sum of probabilities of $M_{r_0}$ members with $ess > minSS$. We consider all `locally optimal' assignments, where probability value cannot be increased without increasing the cost.
\item \textbf{Combining Solutions :} Then, we use dynamic programming to pick one optimal solution per $r_0$ so as to get maximum probability value with the given cost constraint. We discretize costs, say to have granularity $100$. Let $M_{0}, M_{1}, ... M_{D}$ be the sets for which we have found locally optimal solutions. We make $D$ iterations in all, where in the $i^{th}$ iteration, we try different locally optimal solutions for $M_i$ and use it to update the maximum attainable value for each cost. 
\squishend
}
